---
title: "Wisconsin Lake Bathymetry Examples"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{WI_Lake_Bathy_Ex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Setup 

```{r eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
library(Wisclakebathy)
library(dplyr)
library(ggplot2)
library(extrafont)
```

## Introduction

The Wisclakebathy package was designed to help scientists at the Wisconsin DNR
process and analyze lake bathymetry data for Wisconsin lakes. 

<br>

**Obtaining lake bathymetry information.** For many Wisconsin lakes, pdf maps of
lake bathymetry are available [on the WDNR
website](https://dnr.wi.gov/lakes/maps/).

<br>

**Obtaining lake depth vs. lake volume only.** Many of these maps have a lake
depth vs. lake volume curve printed on the map, with the maximum lake depth,
lake volume, and lake area noted in text. If this is the only information
needed, it is relatively fast to use [DataThief](https://datathief.org/) to
extract information from these curves. When using DataThief, first define the
axis, then be sure that when you place the end point marker, the x,y values for
that end point are less than or equal to the printed values for maximum depth
and maximum volume. If not, adjust the axis definition. The functions in this
package can process the raw DataThief results provided the results txt file is
named appropriately and there is an accompanying csv file with the maximum lake
depths and lake volumes noted. See full details about how to organize these
files in the details of the `convert_data_thief_volumes` function (see entry in
the User Manual for this package, or access the same help information by typing
`?convert_data_thief_volumes` with the Wisclakebathy package installed). We do
not recommend estimating lake depth vs. lake area from this lake depth vs. lake
volume curve, as it is difficult to account for irregularly shaped lakes,
particularly when there are multiple basins.

<br>

**Obtaining lake depth, lake area, and lake volume.** If there is no lake depth
vs. lake volume curve provided or if lake depth vs. lake area information is
also desired, it is best to instead use the approach of [Rounds et al.
(2021)](https://conservancy.umn.edu/handle/11299/216182) and use
[ImageJ](https://imagej.nih.gov/ij/download.html) to digitize the proportional
area associated with depth contour intervals. See [their
repository](https://conservancy.umn.edu/handle/11299/216182) for a helpful video
describing this approach. The Wisclakebathy package can convert the depth vs.
proportional areas into actual areas (acres) provided the csv files are named
appropriately and there is an accompanying csv file with the maximum lake area
noted. See full details about how to organize these files in the details of the
`convert_proportion_areas` function (see entry in the User Manual for this
package, or access the same help information by typing
`?convert_proportion_areas` with the Wisclakebathy package installed). Lake
depth vs. lake area information can then be converted to lake depth vs. lake
volumes information using a trapezoidal or conic estimation approch with the
function `calculate_volumes`.

<br>

**Using both approaches.** It is possible that an in-depth analysis of many
lakes will lead to a mix of approaches, with lake depth vs. lake volume
extracted using DataTheif for some lakes but ImageJ used to get lake depth vs.
lake area for other lakes. In this case, it is probably best to maintain one
summary csv file with the WBIC, (maximum) Depth_feet, (maximum) Area_acres, and
(maximum) Volume_acre_ft for each lake that is evaluated using either approach.
Missing information can be denoted in this summary file by (capitalized) NA.
This file can be used as the `max_area_file` in the function
`convert_proportion_areas` and the `maxes_file` in the
`convert_data_thief_volumes`function - the functions will not be confused by the
extra columns. Similarly, DataThief results can be stored in the same directory
as ImageJ results - they will have the same naming pattern
("singlelakename_WBIC"), but different file endings (.txt vs. .csv), so the
functions will not get confused about which files to read.

<br>

## Example 1: Use DataThief to obtain lake depth vs. lake volume

```{r eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
# Note: if knitting straight from vignette, paths must to be relative to
# vignette directory. Otherwise, paths should be relative to project directory.

# Process raw DataThief results
lake_volumes <- convert_data_thief_volumes(path = "../data-raw/CS_DataThief",
                                           max_file = "WI_maxes.csv",
                                           flip = TRUE,
                                           depth_interval = 3)

# Calculate sensitivity of lake volume to changes in depth
vol_loss_10pcnt <- depth_change_given_volume_loss(lake_volumes,
                                                  pcnt_loss = 10)
vol_depth_change_1ft <- volume_loss_given_depth_change(lake_volumes,
                                                       depth_change_ft = 1)

# Combine into summary data frame
volume_changes <- full_join(vol_loss_10pcnt, vol_depth_change_1ft,
                            by = c("WBIC", "lake", "max_depth", "max_vol")) %>%
                  select(WBIC = .data$WBIC, 
                         lake = .data$lake, 
                         max_depth = .data$max_depth, 
                         max_vol = .data$max_vol, 
                         vol_pcnt_loss_1ft_depth = .data$vol_pcnt_loss, 
                         depth_change_10pcnt_vol_loss = .data$depth_change)

# Write out
write.csv(lake_volumes, 
          "../data-raw/CS_Bathymetry.csv", 
          row.names = FALSE)
write.csv(volume_changes, 
          "../data-raw/CS_Bathymetry_volume_changes.csv", 
          row.names = FALSE)
```

## Example 2: Use ImageJ to obtain lake depth, lake area, and lake volume

```{r eval=TRUE, echo=TRUE, warning=FALSE, message=FALSE}
# Note: if knitting straight from vignette, paths must to be relative to
# vignette directory. Otherwise, paths should be relative to project directory.

# Process raw ImageJ results
lake_areas <- convert_proportion_areas(path = "../data-raw/Wisconsin_Hypsography",
                                       max_area_file = "WI_areas.csv",
                                       flip = TRUE)
# Calculate lake volumes using trapezoidal approach (could instead use conical)
lake_volumes          <- calculate_volumes(lake_areas, 
                                           method = "trapezoidal")

# Calculate sensitivity of lake volume and lake area to changes in depth 
vol_loss_10pcnt       <- depth_change_given_volume_loss(lake_volumes,
                                                        pcnt_loss = 10)
vol_depth_change_1ft  <- volume_loss_given_depth_change(lake_volumes,
                                                        depth_change_ft = 1)
area_loss_10pcnt      <- depth_change_given_area_loss(lake_volumes,
                                                      pcnt_loss = 10)
area_depth_change_1ft <- area_loss_given_depth_change(lake_volumes,
                                                      depth_change_ft = 1)

# Combine into summary data frame
volume_changes <- full_join(vol_loss_10pcnt, vol_depth_change_1ft,
                            by = c("WBIC", "lake", "max_depth", "max_vol")) %>%
                  select(WBIC = .data$WBIC, 
                         lake = .data$lake, 
                         max_depth = .data$max_depth, 
                         max_vol = .data$max_vol, 
                         vol_pcnt_loss_1ft_depth = .data$vol_pcnt_loss, 
                         depth_change_10pcnt_vol_loss = .data$depth_change)
area_changes  <- full_join(area_loss_10pcnt, area_depth_change_1ft,
                           by = c("WBIC", "lake", "max_depth", "max_area")) %>%
                 select(WBIC = .data$WBIC, 
                        lake = .data$lake, 
                        max_depth = .data$max_depth, 
                        max_area = .data$max_area, 
                        area_pcnt_loss_1ft_depth = .data$area_pcnt_loss, 
                        depth_change_10pcnt_area_loss = .data$depth_change)
volume_area_changes <- full_join(volume_changes, area_changes,
                                 by = c("WBIC", "lake", "max_depth"))

# To write out:
write.csv(lake_volumes, 
          "../data-raw/WI_Bathymetry.csv", 
          row.names = FALSE)
write.csv(volume_area_changes, 
          "../data-raw/WI_Bathymetry_volume_area_changes.csv", 
          row.names = FALSE)
```

### Visualize Lake Volume

<br>

**Change in Maximum Lake Depth with 10% Volume Loss**
<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
plot_obj <- ggplot(data = volume_area_changes,
                   aes(x = .data$depth_change_10pcnt_vol_loss)) +
            geom_histogram(binwidth = 0.1,
                           position = "identity") +
            labs(x = "Change in Lake Depth (ft)",
                 y = "Number of Lakes") +
            scale_x_continuous(minor_breaks = 0.1,
                               expand = c(0,0)) +
            scale_y_continuous(expand = c(0,0)) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = 12))

plot_obj
```

<br>

**Change in Lake Volume with 1ft Drop in Maximum Lake Depth**
<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
plot_obj <- ggplot(data = volume_area_changes,
                   aes(x = .data$vol_pcnt_loss_1ft_depth)) +
            geom_histogram(binwidth = 1,
                           position = "identity") +
            labs(x = "Change in Lake Volume (%)",
                 y = "Number of Lakes") +
            scale_x_continuous(minor_breaks = 1,
                               expand = c(0,0)) +
            scale_y_continuous(expand = c(0,0)) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = 12))

plot_obj
```

<br>

### Visualize Lake Area

<br>

**Change in Maximum Lake Depth with 10% Area Loss**
<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
plot_obj <- ggplot(data = volume_area_changes,
                   aes(x = .data$depth_change_10pcnt_area_loss)) +
            geom_histogram(binwidth = 0.1,
                           position = "identity") +
            labs(x = "Change in Lake Depth (ft)",
                 y = "Number of Lakes") +
            scale_x_continuous(minor_breaks = 0.1,
                               expand = c(0,0)) +
            scale_y_continuous(expand = c(0,0)) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = 12))

plot_obj
```

<br>

**Change in Lake Area with 1ft Drop in Maximum Lake Depth**
<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
plot_obj <- ggplot(data = volume_area_changes,
                   aes(x = .data$area_pcnt_loss_1ft_depth)) +
            geom_histogram(binwidth = 1,
                           position = "identity") +
            labs(x = "Change in Lake Area (%)",
                 y = "Number of Lakes") +
            scale_x_continuous(minor_breaks = 1,
                               expand = c(0,0)) +
            scale_y_continuous(expand = c(0,0)) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = 12))

plot_obj
```

<br>

### Compare Lake Volume vs. Lake Area Sensitivity

<br>

**Change Maximum Lake Depth with 10% Change in Lake Area vs. Lake Volume**
<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
plot_obj <- ggplot(data = volume_area_changes,
                   aes(x = .data$depth_change_10pcnt_area_loss,
                       y = .data$depth_change_10pcnt_vol_loss)) +
            geom_abline(intercept = 0, 
                        slope = 1,
                        color = "black") +
            geom_point(color = "grey70") +
            labs(x = "Change in Lake Depth with\n10% Change in Lake Area (ft)",
                 y = "Change in Lake Depth with\n10% Change in Lake Volume (ft)") +
            scale_x_continuous(expand = c(0,0)) +
            scale_y_continuous(expand = c(0,0)) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = 12))

plot_obj
```

<br>

**Change in Lake Area vs. Lake Volume with 1ft Drop in Maximum Lake Depth**
<br>
```{r eval=TRUE, echo=FALSE, warning=FALSE, message=FALSE, fig.width=4, fig.height=3}
plot_obj <- ggplot(data = volume_area_changes,
                   aes(x = .data$area_pcnt_loss_1ft_depth,
                       y = .data$vol_pcnt_loss_1ft_depth)) +
            geom_abline(intercept = 0, 
                        slope = 1,
                        color = "black") +
            geom_point(color = "grey70") +
            labs(x = "Change in Lake Area (%)",
                 y = "Change in Lake Volume (%)") +
            scale_x_continuous(expand = c(0,0)) +
            scale_y_continuous(expand = c(0,0)) +
            theme_bw() +
            theme(text = element_text(family = "Segoe UI Semilight",
                                      size = 12))

plot_obj
```
